---
title: "Data Visualization"
author: "Alex Vand"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# General use
library(tidyverse)
library(here)
library(janitor)

# Specifically for dataviz:
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(gapminder)

library(sf)
```

## Get the lizards data

```{r}
lizards <- read_csv(here::here("data_tidy", "lizards.csv"))
```

Histogram (really bad) of lizard weights
```{r}
ggplot(data = lizards, aes(x = weight)) +
  geom_histogram(fill = "orange",
                 color = "purple",
                 size = 1.5,
                 linetype = "dotted")
```

```{r}
ggplot(data = lizards, aes(x = total_length,
                           y = weight)) +
  geom_point(shape = 22,
             fill = "yellow",
             color = "red",
             size = 3,
             alpha = 0.6)
```

Bubble plot where the color of the points changes based on common_name and the size of points varies based on total_length (x-axis).
(Disclaimer: bad idea)

```{r}
ggplot(data = lizards, aes(x = total_length,
                           y = weight)) +
  geom_point(aes(color = common_name,
                 size = total_length),
             shape = 22,
             fill = "black") +
  theme_grey()
```

```{r}
ggplot(data = lizards, aes(x = total_length,
                           y = weight)) +
  geom_point(aes(color = common_name)) +
  theme_grey() +
  facet_wrap(~common_name, ncol = 4)
```

```{r}
ggplot(data = lizards, aes(x = total_length,
                           y = weight)) +
  geom_point(aes(color = common_name)) +
  facet_grid(sex ~ tail)
```

Find total lizard counts by common_name
```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

# Same thing:
# lizard_counts <- lizards %>% dplyr::count(common_name)

lizard_counts_cn_tail <- lizards %>% count(common_name, tail)
```

```{r}
ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, count), x = count)) +
  geom_col(aes(fill = common_name), show.legend = FALSE) +
  labs(x = "Lizard counts")
```

Try converting common name to an ordered factor outside of ggplot.
```{r}
lizard_counts <- lizard_counts %>%
  mutate(common_name = fct_reorder(common_name, count))
```

## Axis scale breaks, limits & labels

Scatterplot: total_length (x) versus weight (y)
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  scale_x_continuous(breaks = c(0, 10, 50, 500),
                     limits = c(0, 500),
                     expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
```

Transform the date column to class Date, then find counts of observatios by date.
```{r}
lizard_counts_date <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  count(date)
```

```{r}
ggplot(data = lizard_counts_date, aes(x = date, y = n)) +
  geom_line() +
  scale_x_date(date_breaks = "3 years",
               date_labels = "%Y") +
  labs(x = "Date",
       y = "Count")
```


```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
#  scale_color_gradientn(colors = c("orange", "red", "purple", "black")) +
  scale_color_stepsn(colors = c("green", "blue", "purple"),
                     breaks = c(0, 20, 40))
```

### Update a color scheme using a palette in paletteer

```{r}
lizards_fct <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))

ggplot(data = lizards_fct, aes(x = total_length, y = common_name)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  labs(x = "Total length (mm)",
       y = "Common Name") +
# View(palettes_d_names)
  scale_fill_paletteer_d(palette = "lisa::AndyWarhol")
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
#   panel.grid = element_blank())
    panel.background = element_rect(color = "purple",
                                    fill = "yellow",
                                    size = 10),
    panel.grid.major.x = element_line(color = "red"),
    panel.grid.major = element_line(color = "blue"),
    panel.grid.minor.y = element_line(color = "orange"),
    plot.background = element_rect(fill = "cyan4"),
    axis.text.x = element_text(color = "orange"),
    axis.title = element_text(color = "yellow", size = 17)
  )
```

## ggprepel

Make a subset from lizards, called ww_lizards, that only contains observations for "western whiptail" lizards from the site "sand"
```{r}
ww_lizards <- lizards %>% 
  filter(common_name == "western whiptail", site == "sand")

ggplot(data = ww_lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_text_repel(aes(label = toe_num), size = 3, max.overlaps = 20)
```

Filter for only one continent and one year, then create a scatterplot of gdpPercap (x) and lifeExp (y), labeled by country name
```{r}
europe <- gapminder %>% 
  filter(continent == "Europe", year == 2007)

ggplot(data = europe, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country), size = 2.5)
```

## gghighlight

```{r}
p <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point()

p + gghighlight(toe_num == 250, label_key = toe_num)
```


```{r}
q <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_line(aes(color = common_name))

q + gghighlight(max(weight) > 30)
```

## patchwork for compound figures

```{r}
(p & q) / q &
  theme_minimal()
```

## A few new graph types

### Marginal plots (ggextra)

```{r}
whiptails <- lizards %>% 
  filter(common_name == "western whiptail") %>% 
  drop_na(total_length, weight)

# unique(whiptails$commonn_name)

ggplot(data = whiptails, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_rug()
```


Marginal plot with boxplots
```{r}
whiptail_plot <- ggplot(data = whiptails, aes(x = total_length, y = weight)) +
  geom_point(aes(color = sex), size = 2) +
  scale_color_manual(values = c("cyan4", "black", "goldenrod"),
                     name = "Sex:",
                     labels = c("female", "juvenile", "male")) +
  theme_minimal() +
  theme(legend.position = "bottom")

ggMarginal(whiptail_plot, type = "boxplot", groupColour = TRUE)
```

### Beeswarm plot
~ jitterplot / histogram / density plot ~
```{r}
ggplot(data = whiptails, aes(x = sex, y = weight)) +
#  geom_violin() +
  geom_beeswarm() +
  geom_boxplot(fill = NA)
  
```

### A heatmap with geom_tile()

Make a new data frame called lizard_counts, starting from lizards, with date converted to class Date. Then count by year and common name the number of lizards observed.
```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date), year = lubridate::year(date)) %>% 
  group_by(year, common_name) %>% 
  summarize(count = n())

# count(year, common_name) this does the same thing as group_by/summarize

class(lizard_counts$date)
```

Plot heatmap:
```{r}
ggplot(data = lizard_counts, aes(x = year, y = common_name)) +
  geom_tile(aes(fill = count)) +
  geom_text(aes(label = count), color = "white", size = 3) +
#  scale_fill_gradientn(colors = c("navy", "red", "orange")) or...
  scale_fill_viridis_c() # colorblind
```

### Make a map!

```{r}
jornada_vegetation <- read_sf(here::here("data_raw", "spatial_vegetation", "doc.kml")) %>%
  select(Name) %>% 
  clean_names()

ggplot(data = jornada_vegetation) +
  geom_sf(aes(fill = name), color = NA) +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys") +
  labs(x = "Longitude", y = "Latitude", fill = "Dominant vegetation:")
```








